### 2013-11-20 results/2013-11-15_Priors_Manuscript/Prior.tex
Manuscript draft for priors on Mj (need revising)

\documentclass[11pt,reqno]{amsart}
\usepackage[top=1in, left=1in, right=1in, bottom=1in]{geometry}                % See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   % ... or a4paper or a5paper or ...
\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
%\usepackage{algorithm, algorithmic}

\usepackage{algorithm}
\usepackage{algpseudocode}

\usepackage{graphicx}

\usepackage{verbatim}
\usepackage{amssymb}
\usepackage{amsmath}

\usepackage{enumitem}

\usepackage{setspace}
\doublespacing

\usepackage{natbib}

%\usepackage{epstopdf}
%\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}

\newcommand{\RR}{I\!\!R} %real numbers
\DeclareMathOperator{\diag}{diag}

\algnewcommand{\Inputs}[1]{%
  \State \textbf{Inputs:}
  \Statex \hspace*{\algorithmicindent}\parbox[t]{.8\linewidth}{\raggedright #1}
}
\algnewcommand{\Initialize}[1]{%
  \State \textbf{Initialize:}
  \Statex \hspace*{\algorithmicindent}\parbox[t]{.8\linewidth}{\raggedright #1}
}

\title[Priors]{Priors for $M_{j}$}
\author{}
%\date{}                                           % Activate to display a given date or no date

\begin{document}

\maketitle

\section{Parametric Gamma Prior}
\subsection{Initialization}
The initial values for the model parameters and latent variables is obtained by a method-of-moments (MoM) procedure. MoM works by setting the population moment equal to the sample moment. We present the initial parameter estimates and provide the derivations here.

Since the distribution is $ M_{j} \thicksim \text{Gamma}(a, b)$. The first and second population moments are

\begin{eqnarray}
	E[M_{j}] =& a b,\\
	E[{M_{j}}^2] =& b^2 a (a+1).
\end{eqnarray}

The first and second sample moments are $m_1 = \frac{1}{J}\sum_{j=1}^J M_{j}$ and $m_2 = \frac{1}{J}\sum_{j=1}^J M_{j}^2$. Setting the population moments equal to the sample moments and we get

\begin{equation}
    a = \frac {{m_1}^2}{{m_2}-{m_1}^2}
\end{equation}

\begin{equation}
    b = \frac {{m_2}-{m_1}^2}{m_1}
\end{equation}

\subsection{Sampling from $p \left( M_{j} |a, b, \theta_{ji},\mu_j\right)$}
The posterior distribution over $M_{j}$ given its Markov blanket is

\begin{equation}
	p( M_{j} |a, b, \theta_{ji},\mu_j) \propto p(\theta_{ji} | \mu_j, M_j) p(M_{j} | a, b)
\end{equation}

We have $ \theta_{ji}\thicksim\text{Beta}(\mu_{j},M_{j})$, and $ M_{j} \thicksim \text{Gamma}(a, b)$. Assuming there is only one replicate,

\begin{equation}
	p( M_{j} |a, b, \theta_{ji},\mu_j) \propto \frac{\Gamma \left({\mu}_{j} {M}_{j} \right)}{\Gamma \left({\mu }_{j}\right)\Gamma \left({M}_{j}\right)} {{\theta}_{j}}^{{\mu}_{j}{M}_{j}-1}{\left(1-\theta\right)_{j}}^{\left(1-{\mu}_{j}\right){M}_{j}-1}\frac{{b}^{a}}{\Gamma \left({a}\right)}M_{j}^{a-1}{e}^{-bM_{j}}
\end{equation}

Since we cannot give an analytical form for this, so we sample from the posterior distribution using the Metropolis-Hastings algorithm.

\section{Jefferys Priors}\label{sec:model_structure}
Jeffreys prior is proposed for invariance by Harold Jeffreys (1946), and defined in terms of the Fisher information. For our ploblem,
\begin{equation}\label{eqn:JefferyPrior}
\pi\left({M}_{j}\right)= I\left({M}_{j}\right)^{\frac{1}{2}}
\end{equation}

Since ${\theta }_{ji}\sim Beta\left( {\mu }_{j},{M}_{j}\right)$, the Fisher information is given by

\begin{equation}\label{equ:JefferyInference}
I\left({M}_{j}\right)={E}_{{M}_{j}}\left[ -\frac{\delta ^{2}\log p\left(\theta _{j}|\mu_{j},M_{j}\right)}{\delta M^{2}_{j}}\right]
\end{equation}

Now to calculate the equations to acquire the Fisher information, we assume there is only one replicate,

\begin{equation}\label{eqn:Betapdf}
p\left({\theta }_{j} \right)= \frac{\Gamma \left({M}_{j} \right)}{\Gamma \left({\mu }_{j} {M}_{j}\right)\Gamma \left(( 1-{\mu }_{j}){M}_{j}\right)} {{\theta}_{j}}^{{\mu}_{j}{M}_{j}-1}{\left(1-\theta\right)_{j}}^{\left(1-{\mu}_{j}\right){M}_{j}-1}
\end{equation}

\begin{equation}\label{equ:JefferyInference1}
\begin{split}
\log p\left(\theta_{j}|\mu_{j},M_{j}\right)& =\log \Gamma \left(M_{j}\right)-\log \Gamma\left(\mu_{j},M_{j}\right)- \log \Gamma\left(1-\mu_{j},M_{j}\right)\\
& + (\mu_{j}M_{j}-1)\log\theta_{j} + ((1-\,u_{j})M_{j}-1)\log(1-\theta_{j})\
\end{split}
\end{equation}

\begin{equation}
\frac{\delta\log p(\theta_{j})}{\delta M_{j}} = \Psi(M_{j}) - \Psi(\mu_{j} M_{j})\mu_{j} - \Psi((1-\mu_{j})M_{j})(1-\mu_{j}) +\mu_{j}\log\theta_{j} + (1-\mu_{j})\log(1-\theta_{j})
\end{equation}

\begin{equation}
\frac{\delta^{2}\log p(\theta_{j})}{\delta M_{j}^{2}}  = \Psi_{1}(M_{j}) - \Psi_{1}(\mu_{j} M_{j})\mu_{j}^{2} - \Psi_{1}((1-\mu_{j})M_{j})(1-\mu_{j})^{2}
\end{equation}

Now we have the Jeffreys' prior for $M_{j}$:

\begin{equation}
\pi\left({M}_{j}\right) = [-\left(\Psi_{1}(M_{j}) - \Psi_{1}(\mu_{j} M_{j})\mu_{j}^{2} - \Psi_{1}((1-\mu_{j})M_{j}){(1-\mu_{j})^{2}}\right)]^{\frac{1}{2}}
\end{equation}

\section{Reference Prior}
We want to make the prior as less informative as possible, so to maximize a certain distance measure of prior and posterior is considered using Kullback-Leibler divergence.

\subsection{KL-Divergence}
For distributions $p$ and $q$, their KL-divergence is defined as $KL\left(p||\right) = E_p[\log\left(p/q\right)] = int_p\left(x\right)\log \frac{p\left(x\right)}{q\left(x\right)}dx$

\subsection{Definition for Reference Prior}
Suppose posterior is $p\left(\theta | x \right)$, depending on pror $p\left(\theta \right)$, marginal is $p\left(x \right)$, then $p\left(\theta \right)$ is called the reference pror if it maximizes $max_{p\left(\theta\right)}E_{p\left(x\right)}KL\left(p\left(\theta|x\right)||p\left(\theta\right)\right)$
The Jeffreys prior is proved to be reference prior in one dimension (by Michael I. Jordan).

\section{Log-normal Prior}
In log-normal distribution, the parameters denoted $\mu$ and $\sigma$, the mean and standard deviation respectively. Log-normal prior was used on $ M_{j}$.

\subsection{Initialization}
The initial values is also obtained by a method-of-moments (MoM) procedure. We present the initial parameter estimates below.

Since the distribution is $ M_{j} \thicksim \text{log-normal}(\mu, \sigma)$. The first and second population moments are

\begin{eqnarray}
	E[M_{j}] = & e^{\mu+\frac{{\sigma}^2}{2}},\\
	E[{M_{j}}^2] = & \left(e^{{\sigma}^2}-1\right)[E[M_{j}]]^2.
\end{eqnarray}

We can acquire the sample moments, then parameters $\mu$ and $\sigma$ can be obtained

\begin{equation}
    \mu = \ln\left(E[M_{j}]\right)-\frac{{\sigma}^2}{2}
\end{equation}

\begin{equation}
    {\sigma}^2 = \ln\left(1+\frac{E[{M_{j}}^2]}{{E[M_{j}]}^2}\right)
\end{equation}

\subsection{Sampling from $p \left( M_{j} |\mu,\sigma, \theta_{ji},\mu_j\right)$}
The posterior distribution over $M_{j}$ given its Markov blanket is

\begin{equation}
	p( M_{j} |\mu, \sigma, \theta_{ji},\mu_j) \propto p(\theta_{ji} | \mu_j, M_j) p(M_{j} | \mu, \sigma)
\end{equation}

We have $ \theta_{ji}\thicksim\text{Beta}(\mu_{j},M_{j})$, and $ M_{j} \thicksim \text{log-normal}(\mu, \sigma)$. Since we cannot give an analytical form for this, so we sample from the posterior distribution using the Metropolis-Hastings algorithm.

\end{document}