### 2015-02-23 doc/bioinformatics2014/revision2/response_to_reviews.txt
Add discussion of sequencing protocol and results.

Reviewer 1

1. When using synthetic DNA to evaluate the sensitivity and sensibility of RVD2, much is dedicated to (currently) unrealistic coverages. For example, in the Figure1 and Table1 (Figure 3) the authors presents a sequence coverage of 5.000x and even 55.000x. Despite being very useful for the own authors as high hypothetical control, I think ~30x - 100x coverage is more informative, and presenting these extremely high coverages can be misleading.

The reviewer correctly points out that much next-generation DNA sequencing data has a median depth in the 30x - 100x range. We have added a paragraph to our discussion of the sensitivity and specificity results (second paragraph in Section 5.2). In that paragraph, we discuss three primary median depth ranges and typical applications for those areas. Briefly, whole genome sequencing typically has depths in the range of 10x - 100x as noted. In addition, targeted cancer gene sequencing typically has median read depths in the range 100x - 1,000x because the region of interest is smaller and there is interest in measuring tumor heterogeneity. Microbial and viral sequencing can have median depths from 1,000x - 100,000x because one is interested in identifying rare subpopulations that may prove resistant to antibiotics or antivirals.

We have chosen a broad range of median depths for our results so that researchers in whole genome sequencing, targeted sequencing and microbial sequencing can see how the methodology would perform in their application area.

2. Despite having the whole genome sequencing for HCC1187 and HCC1187BL cell lines, the authors focus all presented analysis one a single gene. It's not clear why the authors used the coordinates chr7:154738059154782774 (~45Kb), excluding the three first exons (four introns) and the last exon and most of the last intron, when, in fact, PAXIP1 has ~59Kb. No further explanation is given regarding why this specific region was chosen.

Thank you for identifying this important omission. The region of interest for the PAXIP1 gene should have covered the entire 59Kb coding region. We ran RVD2 on the current NCBI start/stop codons (http://www.ncbi.nlm.nih.gov/gene/22976) to give the expanded region of chr7:154,735,400 - 154,794,682. This expanded region yields much better results. In particular, using the shortened region, we identified SNPs at XXX, XXX, XXX, and XXX. An inspection of Fig4 shows that these loci have very similar point estimates of the allele frequency in the case and control. In the analysis with the expanded region, RVD2 estimates a lower across-locus precision parameter (M0) and thus does not call these loci. However, it does call 2 mutations in the full-length PAXIP1 region of interest.

We have updated our results to reflect the accurate start/stop locations for PAXIP1.

3. Instead of using DNA sequencing of primary tumors genomes, which are highly heterogeneous and present many rare (single nucleotide) variants, the authors used the DNA from a commercially available cell line, which is colonially expanded (genetically homogenous). Interestingly, chromosome 7, where the investigated gene is located, is tetraploid on HCC1187[1], therefore, one would expect to find only allele frequencies 0%, 25%, 50%, 75 and 100% on variants located in this chromosome. Figure 5 seems to be in agreement to this hypothesis, except for variants 49704 and 81769, which has marginally lower estimations of allele frequency.Therefore, the variants detected on HCC1187 are not actually rare variants (MAF=25%) and, as shown on Figure 3 (Table1), may be detected by most of currently available methods. Despite the limitations, the analysis is complete, but it would be more informative to apply RVD2 to publicly available (and validated) DNAseq from primary tumors.

4. The authors cite a paper from 2011 to refer for the Synthetic DNA sequencing. Explicit details on the technology, read length and number of reads generated would greatly ease readers understanding of this section.

Section 4.1 describing the details of the experimental protocol for the synthetic sequencing data is indeed brief. We have added the details the reviewer suggested and as expected, the narrative is much clearer. Briefly, Supplementary Table S1 in the Nucleic Acid Research publication shows that each sample has roughly 1,000,000 35bp paired-end reads with no errors in the multiplexing barcodes to give an average read depth of ~100,000 across the synthetic amplicon. We describe this and more detail as suggested in the protocol section.